Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-19 paper/paper.tex
some affiliation politics

\maketitle
\small{
    \textbf{1} INRIA, Parietal team, Saclay, France\\
    \textbf{2} CEA, Neurospin, Gif-sur-Yvette, France\\
    \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
    \textbf{4} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
    \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
    \textbf{6} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
}

@@

\subsubsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy}.
It maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}

@@

Since an action taken in the present moment $t$
might have repercussions in the far future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in evaluation of biological significance,
  whereas
  the NAc is involved in reward evaluation. Both are known to
  be connected to the vmPFC of the DMN.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
% given that there is more uncertainty in the farsighted future.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

>>>>

\maketitle
\small{
    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal team, Saclay, France\\
    \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
    \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
    \textbf{4} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
    \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
    \textbf{6} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
}

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by some kind of \textit{policy}.
It maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}

@@

Since an action taken in the present moment $t$
might have repercussions in the far future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in evaluation of biological significance,
  whereas
  the NAc is involved in reward evaluation. Both are known to
  be connected to the vmPFC of the DMN.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$, in an average sense to be made precise latter. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
% given that there is more uncertainty in the farsighted future.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.
